\begin{abstract} 
Within the fields of High Performance Computing (HPC) and quantum computing, the integration of quantum kernels into HPC
systems has been increasingly explored. Hybrid classical-quantum algorithms, such as the Variational Quantum Eigensolver 
(VQE) utilize the strengths of both paradigms to achieve new grounds within quantum research and overcome modern day hardware limitations. 
This thesis presents the design, development, and benchmarking of an accessible, 
open-source classical-quantum framework that provides a measurable speedup in an MPI distributed and GPU accelerated hybrid simulation for future usage in fields 
such as materials science or complex optimization.
Presently, there is a lack of existing unified middleware that can bridge the low latency QPU control with 
HPC infrastructures, so for that reason this stack extends existing hybrid middlewares to focus on four objectives: designing an 
accessible API to abstract the cloud based QPU backend, MPI distributed parallelization of the Pauli term evaluation classical subroutines, GPU accelerated statevector construction, and a comprehensive performance analysis of the CPU, GPU, and QPU execution paths. The stack's 
evaluation occurs on a single Docker host with a consumer (NVIDIA GTX 1650) GPU and IBM \texttt{ibm\_marrakesh} QPU with the stack achieving a $1.51\times$ speedup in end-to-end wall-clock time 
as when compared to serial and nonoptimized baselines. Results demonstrate GPU acceleration mitigates the CPU scaling bottleneck predicted in 
Amdahl's law from a $0.24\times$ regression at $P=8$ into a $1.51\times$ computational speedup.
Readers will find a complete evaluation of CPU, GPU, and QPU paths, with scaling analysis and stress testing conducted to validate the middleware architecture. 

\textbf{Keywords:} Quantum Systems, Hybrid Systems, Variational Quantum Eigensolver, High-Performance Computing, 
Algorithm Acceleration, MPI, CUDA, Software Middleware.
\end{abstract}

