

\section{Introduction}
    The field of High Performance Computing (HPC) has been undergoing a transformation towards heterogeneous architectures 
    that are not defined by the sole usage of CPUs. In past decades, the integration of General Purpose Graphic 
    Processing Units (GPGPUs) alongside Central Computing Units (CPUs) have defined the limits of this field's 
    computational power. However in recent years, there has been a rising demand for more advanced systems that have the ability to stimulate
    complex molecular dynamics and combinatorial optimization, shining a light on our current classical computational limitations. 
    Issues such as the curse of dimensionality and the exponential scaling of Hilbert space have redirected this field's focus 
    towards the potential advantages of Quantum Processing Units (QPUs).  QPUs have emerged, not to be the replacement for classical computation, but 
    instead as specialized accelerators that will be integrated into our existing classical systems, as these processing units are capable of handing 
    specific computational tasks that are intractable for classical bits ~\cite{rallis2025interfacingquantumcomputingsystems}. However, quantum technology 
    is a new advancement, one that still needs significant refinement and testing due to their current limitations. 
    This means hybrid frameworks are the current industry goal to work towards in the hopes of achieving new ground within the fields of material science, 
    drug discovery, complex optimization, cryptographic analysis, and many others.  

    Currently, we are in the era of Noisy Intermediate-Scale Quantum (NISQ), where quantum hardware is limited and noisy, 
    leaving us with difficulties harnessing the true potential of a "quantum advantage" \cite{harrow2017quantumsupremacy}. The most promising path of achieving this 
    is by the development of coehesive hybrid quantum-classical algorithms ~\cite{rallis2025interfacingquantumcomputingsystems} that combine the computational benefits of each system. 
    Specifically, we are referencing algorithms such as the Variational Quantum Eigensolver (VQE), which contains a feedback loop that 
    a classical optimizer uses to iteratively refine parameters that will then be entered into the connected quantum circuit within 
    the QPU. Sttrategically offloading the computationally heavy classical optimization loop to (heterogeneous) classical 
    resources allows the integrated to QPUs solely focus on the execution of quantum circuits and measuring the resulting expectation values. This aloows developers  
    to bypass the modern day limitations of coherency in QPUs which often arises during the execution of fault-tolerant algorithms such as Shor's algorithm, Grover's algorithm, or Quantum Phase Estimation (QPE).
    Take the QPE algorithm, which requires deep gate depths and long runtimes that inevitably exceed the current temporal limits of  
    qubit coherence, leaving us with decolorant or incorrect results. However, if long term coherency in qubits/quantum hardware was achieved, the QPE algorithm would be 
    able to provide us with an exponential speedup (compared to similar classical methods) when tasked with determining the eigenvalues in large and complex problems ~\cite{peruzzo2014variational}. 
    Classical HPC structures that use MPI, CUDA and multi-node CPU/GPU clusters, excel with tasks that handle dense parallelized numerical workloads and larger 
    scale simulations. However these same classical systems will struggle with simulations where the dimension of the problem grows exponentially alongside 
    the size of the physical system (such as electrons within molecules) being observed. This challenge occurs specifically with calculating the arrnagement and behavior of electrons in a molecule, but can be addressed 
    through the usage of hybrid systems that leave the quantum component to map the system's wavefunction directly onto the qubits while the classical 
    components handle the refinement of the variational parameters that will be inputted into quantum circuits \cite{rallis2025interfacingquantumcomputingsystems}.
    
    Despite the theoretical potential of hybrid collaboration, this field currently has a significant architectural gap, as  
    there is a lack of specialized middleware that bridges quantum and classical components. Classical HPC frameworks (ones that are optimized 
    for MPI and CUDA) are not designed to communicate directly with quantum control stacks or hardware. This requires quantum researchers and developers to manually handle the complex 
    data serialization and the frequent network latencies that occur in this communication. Existing quantum programming frameworks, such as Qiskit and Cirq, 
    excel with circuit-level QPU control but lack features to effectively integrate with distributed HPC environments, lacking the ability to seamlessly  
    communicate between the two systems. Without the existence of a standardized hybrid middleware, quantum researchers are facing issues such as high tail latencies that are observed when synchronizing
    both HPC nodes and cloud-based QPUs, leading to performances bottlenecks that ruins the possibility of a speedup offered by a
    hybrid system. This issue, combined with the complex task of data serialization for quantum hardware and the potential lack of understanding of the intricate communication  
    between quantum and classical systems, leaves a difficult barrier for researchers to overcome. 

    The proposed project that this paper will undertake is the development of a hybrid quantum-classical software stack that builds the middleware layer needed
    to abstract hardwares and automate the dispatching of tasks. This software stack will design a clear, intuitive API that will abstract the 
    backend of the cloud-based QPU, removing the need for quantum researchers or developers to manually handle the communication complexities in a hybrid system. This hybrid stack 
    will include automated dispatching logic, HPC connectors to the Qiskit framework, and classical kernels 
    that use MPI and CUDA to parallelize the classical optimization subroutines for an added speedup. The stack is benchmarked against VQE workflows 
    on multi-node GPU clusters, aiming to demonstrate measurable progress towards the "quantum advantage," with a 2-3x 
    speedup in the end-to-end wall-clock time when being compared to the baseline single-node and non-optimized classical implementations. Ultimately, 
    the purpose of this work is the design, implementation, and benchmarking an accesible, open-source HPC software framework that achieves a measurable 
    workflow acceleration through integrating QPUs with multi-node classical resources. 





